The zero--one laws of Kolmogorov and Hewitt--Savage in categorical probability

### 2019-12-04 categorical-zero-one.tex
take out todo remark about spaces
removed wrong proof, added counterexample

\todo[inline]{E: Not sure these examples (about top. spaces) are that great.. it seems like we do an inordinate amount of work to get the infinite tensor products, compared to the results. What do you think?}

@@

We can therefore instantiate our \Cref{thm:kolmog,thm:hewsav} in $\Kl(H)$, which gives the following results as special cases.

\begin{corollary}
    Let $(X_i)_{i \in J}$ be a family of topological spaces, $Y$ a Hausdorff space, and let $f: \prod_i X_i \to Y$ be a continuous function
    which is independent of any finite subset of the input.
    Then $f$ is constant.
    \label{vietoris_kolmogorov}
\end{corollary}
\begin{corollary}
	Let $X$ be a topological space, $Y$ a Hausdorff space, $J$ an infinite set, and let $f: \prod_{i\in J} X \to Y$ be a continuous function which is invariant under any finite permutation of the inputs.
    Then $f$ is constant.
    \label{vietoris_hewitt_savage}
\end{corollary}
\begin{proof}[Proof of both]
	These results follow from \Cref{thm:kolmog,thm:hewsav} applied to $\Kl(H)$,
	using the morphism $p : I \to X_J$ corresponding to the trivial closed subset $X_J \subseteq X_J$.
	The conclusion is that the closure of its image, call it $A$, satisfies
	\[
		\cl(\{(a,a) \mid a \in A\}) = A \times A
	\]
	as subsets of $Y \times Y$.
    But since the diagonal is closed in $Y$, this clearly implies that $A$ must be a singleton.
\end{proof}

>>>>

@@

We can therefore instantiate our \Cref{thm:kolmog} in $\Kl(H)$, which gives the following result as a special case.

\begin{corollary}
    Let $(X_i)_{i \in J}$ be a family of topological spaces, $Y$ a Hausdorff space, and let $f: \prod_i X_i \to Y$ be a continuous function
    which is independent of any finite subset of the input.
    Then $f$ is constant.
    \label{vietoris_kolmogorov}
\end{corollary}
\begin{proof}
	These results follow from \Cref{thm:kolmog} applied to $\Kl(H)$,
    using the morphism $p : I \to X_J$ corresponding to the trivial closed subset $X_J \subseteq X_J$.
	The conclusion is that the closure of its image, call it $A$, satisfies
	\[
		\cl(\{(a,a) \mid a \in A\}) = A \times A
	\]
	as subsets of $Y \times Y$.
    But since the diagonal is closed in $Y$, this clearly implies that $A$ must be a singleton.
\end{proof}

However, we cannot instantiate the Hewitt-Savage 0-1 law, because $\Kl(H)$ is not in fact causal, as the following example shows.
\begin{example}
    Let $A = \{*\}, X = Y = Z = [0,1]$.
    Let $f: A \to HX$ send $*$ to $X$, and define
\begin{equation}
g(x) =  \begin{cases} [0,1/2] & x < \frac{1}{3}\\ \{1/2\} & \frac{1}{3} < x < \frac{2}{3}\\ [1/2,1] & x > \frac{1}{2}\end{cases}
\end{equation}
\begin{equation}
h_1(y) = [0,1]
\end{equation}
\begin{equation}
h_2(y) = \begin{cases}[0,1] & y \neq \frac{1}{2}\\ \{0\} &  y = \frac{1}{2}\end{cases}\end{equation}

One can verify that these are indeed continuous maps.
To check the hypothesis of the statement of causality, we must verify that two maps $A \to Y \times Z$ agree - these are just closed subsets of $[0,1]^2$
From the definition of the composition, we see that any point $(y,z)$ with $y \in g(x)$ for some $x$ and $z \in h_1(y)$ must lie in the set corresponding to $h_1$,
and similarly for $h_2$.
Hence the set corresponding to $h_1$ contains all points, while the set corresponding to $h_2$ contains all points with $y \neq 1/2$.
But the closure of $\{(y,z) \mid y \neq 1/2}$ in $[0,1]^2$ is everything, so both sets must equal the whole space.

Hence the hypothesis is satisfied.
But one can verify that the point $(x,y,z) = (1/2,1/2,1)$ is in the subset of $X \times Y \times Z$ corresponding to $h_1$, but not $h_2$.
\end{example}
